Choosing an appropriate scalar transformation to normalise wpd

### 2021-02-04 scalar_transformation.Rmd
transformation with my earlier attempt included

 - How the distribution of the raw wpd (without any transformation for normalization) looks across nfacets and nx? Both shape and scale of the distribution changes for different nx and nfacet categories.

@@

```{r raw}
N01 %>% 
  ggplot(aes(x = value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  scale_x_continuous(breaks = scales::breaks_extended(3)) + 
  xlab("raw wpd")
```

 - Plot the values of raw wpd against nx*nfacet to see the rough relationship

```{r}
N01 %>% 
  ggplot(aes(x=nx*nfacet, y = value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

@@

- Fit a log-linear relationship of median(values) to log(nx*nfacet) as discussed today. The fit is pretty good.

@@

- Make it a horizontal line so that the values are not affected by 
nx\*nfacet by plotting (value - intercept)/log(nx\*nfacet) against log(nx*nfacet). The line actually becomes horizontal after this transformation.

```{r}
intercept <- fit_lm2$coefficients[1]
slope <- fit_lm2$coefficients[2]

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept)/log(nx*nfacet))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")

```
- See distribution of the transformed variable across nx and nfacets. For higher values of nx and nfacet, the distribution is similar, but not for lower values.

```{r}
N01 %>% 
  ggplot(aes(x = (value - intercept)/(log(nx*nfacet)))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd") +
  scale_x_continuous(breaks = scales::breaks_extended(3))

@@

```{r}

@@

 N01 %>% 
  ggplot(aes(x = log(sqrt(nx*nfacet))/value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd norm")
```

@@

```{r}
fit2 <- lm(log(sqrt(nx*nfacet))/value ~ log(nx*nfacet), N01)
intercept <- fit2$coefficients[1]
slope <- fit2$coefficients[2]

N01 %>% 
ggplot(aes(x = 
             (log(sqrt(nx*nfacet)) - intercept)/log(nx*nfacet))) +
geom_density(fill = "blue") +
facet_grid(nx~nfacet,
labeller = "label_both") +
xlab("wpd norm")

>>>>

 - How the distribution of the raw wpd (without any transformation for normalization) looks across nfacets and nx? Both shape and scale of the distribution changes for different nx and nfacet categories as could be seen in Figure \ref{fig:raw}.

@@

```{r raw, fig.cap="Distribution of raw wpd is plotted across different nx and nfacet categories. Both shape and scale of the distribution changes for different nx and nfacet categories."}
N01 %>% 
  ggplot(aes(x = value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  scale_x_continuous(breaks = scales::breaks_extended(3)) + 
  xlab("raw wpd")
```

 - Plot the values of raw wpd against nx*nfacet to see the rough relationship (\ref{fig:quadratic}).

```{r quadratic, fig.cap = "The raw wpd is plotted against nx*nfacet and the blue line represents the median of the multiple values for each nx*nfacet levels."}
N01 %>% 
  ggplot(aes(x=nx*nfacet, y = value)) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

@@

- Fit a log-linear model of median(values) to log(nx*nfacet) as discussed today. The fit is pretty good and the summary of the fit is shown below.

@@

- Make it a horizontal line so that the values are not affected by 
nx\*nfacet by plotting $wpd_{norm} = (value - intercept)/log(nx*nfacet)$ against log(nx*nfacet). The line actually becomes horizontal after this transformation as could be seen in Figure \ref{fig:horizontal}.

```{r horizontal, fig.cap = "wpd_{norm} is plotted against log(nx*nfacet). The relationship changed from being non-linear to almost horizontal. The horizontal line would imply that median values of wpd are unaffected by the value of nx* nfacet."}

intercept <- fit_lm2$coefficients[1]
slope <- fit_lm2$coefficients[2]

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept)/log(nx*nfacet))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")  +
  ylab("wpd_norm = (wpd - intercept)/log(nx*nfacet)")

```
- See distribution of the transformed variable $wpd_{norm}$ across nx and nfacets. For higher values of nx and nfacet, the distribution is similar, but not for lower values.

```{r, fig.cap = "The distribution of the transformed wpd is plotted. The shape and the scale of the distribution still changes across different nx and nfacet levels, but seems uniform for higher levels." }
N01 %>% 
  ggplot(aes(x = (value - intercept)/(log(nx*nfacet)))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd") +
  scale_x_continuous(breaks = scales::breaks_extended(3))

@@

```{r, fig.cap = "Earlier attempt: A different transformation was used for making the relationship linear and then horizontal. The linear relationship looks promising but when the same method was applied to make it horizonal and eventually look at the distribution, it was again found that the distributions across different nx and nfacets are not similar."}

@@

 N01 %>% 
  ggplot(aes(x = log(sqrt(nx*nfacet))/value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("log(sqrt(nx*nfacet))/value")
```

@@

```{r, fig.cap="Same approach was employed to make it" }
fit2 <- lm(log(sqrt(nx*nfacet))/value ~ log(nx*nfacet), N01)
intercept <- fit2$coefficients[1]
slope <- fit2$coefficients[2]

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), 
             y = ((log(sqrt(nx*nfacet))/value - intercept)/log(nx*nfacet)))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")

N01 %>% 
ggplot(aes(x = 
             (log(sqrt(nx*nfacet))/value - intercept)/log(nx*nfacet))) +
geom_density(fill = "blue") +
facet_grid(nx~nfacet,
labeller = "label_both") +
xlab("wpd norm")
